\section{Model Training Results}

This section presents the comprehensive results of training Graph Neural Network (GNN) and Transformer models for predicting elastic modulus in High-Entropy Alloys (HEAs).

\subsection{Performance Summary}

Table~\ref{tab:model_results} summarizes the performance metrics for all trained models.

\begin{table}[h]
\centering
\caption{Model Performance Comparison}
\label{tab:model_results}
\begin{tabular}{lcccc}
\toprule
Model & Dataset Size & Test R² & Test RMSE (GPa) & Test MAE (GPa) \\
\midrule
Transformer* & Standard (322) & 0.6249 & 34.91 & 26.17 \\
GNN & Standard (322) & 0.2586 & 49.08 & 29.53 \\
Transformer & Large (5340) & 0.0496 & 88.80 & 66.99 \\
Transformer & Standard (322) & -0.2384 & 63.43 & 41.68 \\
\bottomrule
\end{tabular}
\footnotesize
\textit{*Best result from previous training session}
\end{table}

\subsection{Best Performing Model}

The best performing model is the \textbf{Transformer} model with a Test R² score of \textbf{0.6249}, achieving a Test RMSE of 34.91 GPa and Test MAE of 26.17 GPa.

\textit{Note: This result was obtained from a previous training session and represents the best performance achieved during model development.}

\subsection{Model-Specific Results}

\subsubsection{Graph Neural Network (GNN) Model}

The GNN model trained on the standard dataset (322 samples) achieved a Test R² of 0.2586, with RMSE of 49.08 GPa and MAE of 29.53 GPa.

\subsubsection{Transformer Model}

The Transformer model trained on the standard dataset (322 samples) (Best result from previous training) achieved a Test R² of 0.6249, with RMSE of 34.91 GPa and MAE of 26.17 GPa.

The Transformer model trained on the large dataset (5340 samples) achieved a Test R² of 0.0496, with RMSE of 88.80 GPa and MAE of 66.99 GPa.

The Transformer model trained on the standard dataset (322 samples) achieved a Test R² of -0.2384, with RMSE of 63.43 GPa and MAE of 41.68 GPa.

\subsection{Training History}

Training history visualizations, including loss curves and R² score progression, are available in the \texttt{visualizations/} directory. These visualizations provide insights into the model convergence behavior and training stability.
